A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

\begin{figure}[t]
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.65cm]{images/projim.png}}}
    \quad
    \subfloat[\centering Projection of original image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.65cm]{images/projim_warped_scaled.png}}}
    \quad
    \subfloat[\centering Projection of our warped image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach. (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face towards the scene plane, a camera we shall in what follows refer to as `downwards facing', for brevity. To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera. Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the desired principal axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane. We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes to match those of the virtual camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely. Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely. The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector---for the recovered projector given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected.  We carry out this placement according to a small handful of steps. It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes. It is, however, only if the virtual projector be placed to face the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face downwards to the scene plane precisely. Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
\begin{equation}
h = w \times f_\text{proj} / w_\text{im},
\end{equation}
where $f_\text{proj}$ [px] is the focal length of the projector. These three steps are illustrated in Figure~\ref{fig:virtualproj}. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera. %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}

@@

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}

>>>>

\begin{figure}[t]
    \centering
    \subfloat[\centering Original image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection of original image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly. (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Our warped image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection of our warped image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach. (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Notably, we do this in a manner placing the $X$- and $Y$-axes of the augmentation in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane (cf.\ Figure~\ref{fig:eval}(a)). In what follows,  \textit{we shall refer to this camera as `downwards facing'} for brevity, in quotes since the pose of the camera is likely to deviate at least gently from facing downwards to the scene plane directly (an effect for which we shall carry out a correction). Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the $X$- and $Y$-axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane. We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes to match those of the virtual camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

%This makes setting those axes particularly easy from the user viewpoint, since it reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane. Because we wish to set the principal axes of our augmentations in terms of a rotation about the normal vector of the scene plane,  To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera.

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely. Accordingly, we rotate the recovered `downwards-facing' camera (frustum in black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely. The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions. Note in (b) that the virtual projector's principal axes are aigned with those of the virtual camera in Figure~\ref{fig:virtualcam}(b).} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect correcting for projective distortions, placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and setting the dimensions of the augmentation in line with the desired target dimensions.  We carry out this placement according to a small handful of steps. It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes. It is, however, only if the virtual projector be placed to face the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face downwards to the scene plane precisely. Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
\begin{equation}
h = w \times f_\text{proj} / w_\text{im},
\end{equation}
where $f_\text{proj}$ [px] is the focal length of the projector. These three steps are illustrated in Figure~\ref{fig:virtualproj}. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera. %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}

@@

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane, here the left view of a stereo camera).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}
